A comprehensive framework to capture the arcana of neuroimaging analysis

### 2018-10-12 manuscript/manuscript.tex
more citation insertion

Despite the availability of well-established neuroimaging analysis
packages (\cite{cox_afni:_1996}; \cite{smith_advances_2004}; \cite{friston_statistical_2007};
Tournier et. al. 2012), the arcana of neuroimaging analysis is substantial due to
the range of available tools, tuneable parameters, and imaging sequences
involved (\cite{cusack_automatic_2015}). The distribution of complete
``end-to-end'' workflows, from acquired data to publication results, is
necessary for routine reproduction because of the effort required to
accurately exactly reimplement such analyses (\cite{kennedy_neuroimaging_2018}). It is also
difficult to adapt existing workflows to new studies without detailed
knowledge of their design. Therefore, flexible, portable and complete
workflows are important to promote reproduction and code reuse in
neuroimaging research.

A barrier to designing portable and complete workflows is the
heterogeneity of data storage conventions (\cite{marcus_extensible_2007}; Das et.
al. 2011; \cite{gorgolewski_brain_2016}). To address this, the emerging Brain
Imaging Data Standard (BIDS) (\cite{gorgolewski_brain_2016}) provides a way
to standardise the storage of neuroimaging data in file system
directories. BIDS specifies strict file and directory naming
conventions, which facilitate the design of portable \emph{BIDS Apps}
(https://bids-apps.neuroimaging.io). However, for research groups with
sufficient informatics support, software-managed repositories (Marcus et. al. 2007; Das et. al. 2011) can provide additional features, such as
flexible access-control and automated pipelines. Therefore, for
published workflows, the choice of repository should be transparent in
order to maximise their audience.

While neuroimaging analyses are generally amenable to standardisation
(\cite{kennedy_neuroimaging_2018}), minor modifications are often required to accommodate
idiosyncrasies of the acquisition protocols in use at different sites
(Esteban et al. 2018). Therefore, workflows may require conditional
logic in construction or execution to be portable. Nipype is a flexible
Python framework for neuroimaging analysis in which workflows are
constructed programmatically in Python (Gorgolewski et al. 2011).
Programmatic construction allows for rich control-flow logic that is not
readily available in alternative workflow frameworks (Cusack et. al. 2015; \cite{achterberg_fastr:_2016}; \cite{amstutz_common_2016}), and has been used
to implement workflows that are robust to differences in fMRI protocols
across a large number of sites (Esteban et al. 2018).

@@

Similar to data specifications, parameter specifications are included in
the Study class by providing a list of \emph{ParameterSpec} objects to
the \emph{add\_parameter\_specs} class attribute. ParameterSpec objects
are initialised with a name and default value. Special parameters that
specify a qualitative change in the analysis, for example using ANTs
registration (\cite{avants_reproducible_2011}) instead of FSL registration (Smith et. al. 2004), are specified by the \emph{SwitchSpec} subtype.
SwitchSpecs take a name, default value and a list of accepted values at
initialisation.

>>>>

Despite the availability of well-established neuroimaging analysis
packages (\cite{cox_afni:_1996}; \cite{smith_advances_2004}; \cite{friston_statistical_2007};
\cite{tournier_mrtrix:_2012}), the arcana of neuroimaging analysis is substantial due to
the range of available tools, tuneable parameters, and imaging sequences
involved (\cite{cusack_automatic_2015}). The distribution of complete
``end-to-end'' workflows, from acquired data to publication results, is
necessary for routine reproduction because of the effort required to
accurately exactly reimplement such analyses (\cite{kennedy_neuroimaging_2018}). It is also
difficult to adapt existing workflows to new studies without detailed
knowledge of their design. Therefore, flexible, portable and complete
workflows are important to promote reproduction and code reuse in
neuroimaging research.

A barrier to designing portable and complete workflows is the
heterogeneity of data storage conventions (\cite{marcus_extensible_2007}; Das et.
al. 2011; \cite{gorgolewski_brain_2016}). To address this, the emerging Brain
Imaging Data Standard (BIDS) (\cite{gorgolewski_brain_2016}) provides a way
to standardise the storage of neuroimaging data in file system
directories. BIDS specifies strict file and directory naming
conventions, which facilitate the design of portable \emph{BIDS Apps}
(https://bids-apps.neuroimaging.io). However, for research groups with
sufficient informatics support, software-managed repositories (\cite{marcus_extensible_2007}; Das et. al. 2011) can provide additional features, such as
flexible access-control and automated pipelines. Therefore, for
published workflows, the choice of repository should be transparent in
order to maximise their audience.

While neuroimaging analyses are generally amenable to standardisation
(\cite{kennedy_neuroimaging_2018}), minor modifications are often required to accommodate
idiosyncrasies of the acquisition protocols in use at different sites
(Esteban et al. 2018). Therefore, workflows may require conditional
logic in construction or execution to be portable. Nipype is a flexible
Python framework for neuroimaging analysis in which workflows are
constructed programmatically in Python (Gorgolewski et al. 2011).
Programmatic construction allows for rich control-flow logic that is not
readily available in alternative workflow frameworks (\cite{cusack_automatic_2015}; \cite{achterberg_fastr:_2016}; \cite{amstutz_common_2016}), and has been used
to implement workflows that are robust to differences in fMRI protocols
across a large number of sites (Esteban et al. 2018).

@@

Similar to data specifications, parameter specifications are included in
the Study class by providing a list of \emph{ParameterSpec} objects to
the \emph{add\_parameter\_specs} class attribute. ParameterSpec objects
are initialised with a name and default value. Special parameters that
specify a qualitative change in the analysis, for example using ANTs
registration (\cite{avants_reproducible_2011}) instead of FSL registration (\cite{smith_advances_2004}), are specified by the \emph{SwitchSpec} subtype.
SwitchSpecs take a name, default value and a list of accepted values at
initialisation.
